State of The Practice in Indirect Conflicts: Work Flows, Tools, and Future Direction

### 2014-03-10 iced-2013/paper.tex
ICED bird's edits.

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of researchers, \textit{direct} and
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and
positive developer feedback. Indirect conflicts (ICs) are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influence negatively. Indirect conflict tools however, have
not shared the same success as direct conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, which causes tools to eventually be
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue.
Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence to leverage developers in indirect conflict tools.
However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
problems as their evaluations have
been small scale, involving a very limited number of industrial participants or student evaluators.

We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
To do so, we take a step back and intend to obtain a broader view of indirect conflicts
with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts to better understand why indirect conflict tools fail to achieve
the success of other domain tools. We determine: general events which 
cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
use, and what practitioners want from new tools.
We ask:

@@

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that:
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results,
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts.

@@

Not all participants had strong opinions or any experience on every category listed above. For these participants, answers
to the specific categories were not required. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

@@

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
theory techniques for analysis and not for a final theory). For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding.
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
3 criteria that are commonly used in evaluating grounded theory studies.

@@

A large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract said 12 developers. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark.
In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contacts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

@@

In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration)
as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration
tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project
themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run
time testing. The questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.

Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time
frames of when events occurred. Questionnaire data show that 23\% of developers use native IDE tools and 21\% said they use
features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading
of formal documentation.

%Very important
Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
approach. This means that while developers do have processes and tool for prevention, they would rather spend their time
at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.

@@

Having reported the aforementioned trends in industry (what, when, how, mitigation, and resolution), we now discuss these
results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
false positives, and scalability.
As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
industry settings.

\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
false positives, as the changes being reported are not causing them to take action. These reported non actionable responses
lead to information overload and false positives.}

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore
information being presented to them, rendering the system useless. These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload.
Developers simply do not want to know about events which effect them, but require no action on their part.

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption.
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible.
In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers
tend to use the bare amount of communication
towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
Bolici et. al~\cite{Bolici:2009} also give the
possibility of developer stigmergy as the reason that developers do not explicitly communicate with each other about issues opposed to
time management. The idea
that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication
and unwanted awareness.
\textit{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
influence of total, or near total
awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
with higher quality or understanding of the product.
A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to
producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
be more productive.

\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.
Focusing on cure is a direct response to the issue of scalability found in indirect conflict tools. Trying to prevent all conflicts 
is too large a problem to handle, while curing existing conflicts is manageable by a developer.}

We have seen through the results, that developers posses both tools and practices for the prevention,
detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
accomplished essentially by
not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
developers are spending little to no time in prevention. Developers do however spend a large amount of time in detection and cure through the
writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
conflicts occurred at either the detection or cure levels. ``You're reaction time is much more crucial''
said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
matter what prevention was taken, the issue must simply be resolves as quick as possible now.

The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts.
Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
presented in this area is if curative measures are more productive than preventative measures.

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
in that they result in a trade off
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

A last interesting observation about current curative measures being taken in industry, is that developers view testing of software
as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
However, if a test never fails, could it not be said that it is preventing changes from causing it to fail rather than detecting
when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
``prevention is the goal'' was commonly said by developers.

This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
designers, architects, or managerial stakeholders.
From the project manager's interviewed it was shown that they are heavily favored towards planning and
prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
is being viewed in research.

%mature point, static analysis
\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
some software project configurations and software languages to be quite prone to indirect conflicts. This is yet another issue related
to scalability of existing indirect conflict tools.}

@@

This added level of context, through the notion of a mature point,
would add to the prevention versus cure debate as previously discussed. Prevention may be a better
choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
and the mature point in order to fully understand the nature of indirect conflicts and their context.

While this added sense of understanding is important to indirect conflicts, contextual attributes of a project's current progress
or measures of performance are often more synonymous with project management
rather than software developers themselves. An understanding of a project's evolution, a mature point being reached as an example,
may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
organizational structures of software teams. Between increased modularization (multiple sub projects or repositories),
cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have
become quite limited in identifying and tracking dependencies where they were once strong.
In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
ill fit of current static analysis to current development practices..

@@

\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly
produce extremely complex answers.}

@@

\textbf{Give Developers What They Want} \textit{Developers have their own notions of what they want, and how ``good''
that something is for their productivity. We should pay attention to what they want, but also be mindful that what
developers want may not always be good for them.}

@@

\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
multiple languages, and often having a database schema, we should focus on better techniques for static analysis based
on what industry standards are for languages and project structures.}

>>>>

As Software Configuration Management (SCM) has grown, the maturity and norm of parallel
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of researchers, \textit{direct} and
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and
positive developer feedback. Indirect conflicts (ICs) are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influenced negatively. Indirect conflict tools however, have
not shared the same success as direct conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, which causes tools to eventually be
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue.
Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
such as Cataldo et al.'s~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence.
However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
problems as their evaluations have
been small scale, involving a limited number of industrial participants or student evaluators.

We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
To do so, step back and intend to obtain a broader view of indirect conflicts
with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
scalability issues to better understand why indirect conflict tools fail to achieve
the success of other domain tools. We determine: general events which 
cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
use, and what practitioners want from new tools.
We ask:

@@

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that:
\begin{itemize}
	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding
	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
	\item developers do not want awareness mechanisms which provide non actionable results
	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
				context of indirect conflicts
\end{itemize}

@@

Not all participants had strong opinions or any experience on every category listed above. For these participants, answers
to the specific categories were not required. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being part of any one software project for extended periods of time.

@@

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
theory techniques for analysis and not for a final theory). For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding (relating
code and concepts to each other).
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
3 criteria that are commonly used in evaluating grounded theory studies.

@@

A large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract said 12 developers. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark.
In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contracts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

@@

In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration)
as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration
tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. In corroboration, 
questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.

Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time
frames of when events occurred. Questionnaire data show that 23\% of developers use native IDE tools and 21\% said they use
features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
mentioned a form of communication with other developers in aid to resolving these conflicts and only 4\% mentioned the reading
of formal documentation.

%Very important
Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
approach. This means that while developers do have processes and tools for prevention, they would rather spend their time
at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.

@@

We now discuss the aforementioned
results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
false positives, and scalability.
As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
industry settings.

\subsection{Unwanted Awareness}

\textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
false positives, as the changes being reported are not causing them to take action.}

As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore
information being presented to them, rendering the system useless.

\begin{itemize}
	\item These false positives are caused by a difference in what
				developers consider to be false positives versus what awareness literature considers they are.
\end{itemize}

This disjoint, as will
be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action.

\begin{itemize}
	\item Developers don't care about what they don't need to account for
\end{itemize}

``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload.
Developers simply do not want to know about events which effect them, but require no action on their part.

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption.
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible.
In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers
tend to use the bare amount of communication
towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
\textit{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
influence of total, or near total
awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
with higher quality or understanding of the product.
A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to
producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
however, when given the choice, developers tend to lean towards less awareness in order to be more productive in their eyes.

\subsection{Prevention versus Cure}

\textit{Developers would rather spend their time on curative measures (fixing problems after they
arise) than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.
Focusing on cure is a direct response to the issue of scalability found in indirect conflict tools. Trying to prevent all conflicts 
is too large a problem to handle, while curing existing conflicts is manageable by a developer.}

We have seen through the results, that developers possess both tools and practices for the prevention,
detection, and resolution of indirect conflicts.

\begin{itemize}
	\item Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.
\end{itemize}

This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop potential or some probability of issues, or indirect conflicts, form occurring in the first
place, while cure refers to fixing non probable (have happened) issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
accomplished essentially by
not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
developers are spending little to no time in prevention (or what the perceive as prevention). Developers do however spend a 
large amount of time in detection and cure through the
writing of tests and the debugging of issues (although writing tests can be viewed as prevention while their utility is only see at detection). 
In fact, most improvements mentioned by developers in regard to dealing with indirect
conflicts occurred at either the detection or cure levels. ``You're reaction time is much more crucial''
said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
matter what prevention was taken, the issue must simply be resolves as quick as possible now.

The lack of prevention process and tools being used is believed to be due to 2 factors as discovered in the interview process. 
The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts.
Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
presented in this area is if curative measures are more productive than preventative measures.

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
in that they result in a trade off
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be thinking of how to move developers into a more preventative mindset. If the inverse is true, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

A last interesting observation about current curative measures being taken in industry, is that developers view testing of software
as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
This may suggest that if, given the right tools, developers may no longer view preventative
measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
``prevention is the goal'' was commonly said by developers.

This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
designers, architects, or managerial stakeholders.
From the project manager's interviewed it was shown that they are heavily favored towards planning and
prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
approaches to their developers.

%mature point, static analysis
\subsection{Gap in Software Evolution Analysis}

\textit{Due to a lack of productive software analysis tools, caused by project infrastructures
and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
some software project configurations and software languages to be quite prone to indirect conflicts. This is yet another issue related
to scalability of existing indirect conflict tools.}

@@

This added level of context,
would add to the prevention versus cure debate as previously discussed. Prevention may be a better
choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
causes more bugs than can be prevented. It may be imperative to discover more on project stability
and the mature point in order to fully understand the nature of indirect conflicts and their context.

While this added sense of understanding is important to indirect conflicts, an understanding of a project's evolution, 
a mature point being reached as an example,
may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
organizational structures of software teams. Between increased modularization (multiple sub projects or repositories),
cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
become quite limited in identifying and tracking dependencies where they were once strong.
In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
ill fit of current static analysis to current development practices..

@@

\textbf{Prevention versus Cure} \textit{The need for continued study of the open question
``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly
produce extremely complex answers.}

@@

\textbf{Give Developers What They Want} \textit{Developers have their own notions of ``good''
that something is for their productivity. We should pay attention to what they want, but also be mindful that what
developers want may not always be good for them.}

@@

\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
multiple languages, and more modular structures, we should focus on better techniques for static analysis based
on what industry standards.}
